---
title: "CIRCA CHEM"
author: "Stavros Oikonomou"
date: "11/9/2022"
output:
  html_document: default
  word_document: default
---

```{r, warning=F,message=F, include=F}
rm(list = ls(all=TRUE))
ipak <- function(pkg){
  new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
  if (length(new.pkg)) 
    install.packages(new.pkg, dependencies = TRUE)
  sapply(pkg, require, character.only = TRUE)
}

# usage
packages <- c("caret","dplyr","ggplot2","lubridate","readxl", "stringr",
              "kableExtra","tableone","knitr","xlsx","magrittr","data.table")

ipak(packages)

rm(ipak,packages)
```

```{r, include=FALSE, warning=FALSE, message=FALSE}
#load the data
df <- read.csv("./data_files/CIRCACHEMStudy_DATA_2021-03-04_1440_3.csv") 

#descriptive data
df_descr <- read_excel("./data_files/Copy of CIRCA CHEM_Participants Registry_29.12.20_2.xlsx") %>%
  janitor::clean_names()
```


```{r}
setnames(df,"ï..record_id","record_id")

#excluding the 4 participants. 9 and 11 leave the second week and 13 and 34 they never start.
excluded_patients <- c(9,11,13,34)

#list of participants in group A.
groupA <- c(1,2,5,6,10,14,17,18,21,22,25,
            26,29,30,33,35,37,38,41,42,45)

df <- df %>%
  filter(!record_id %in% excluded_patients) %>%
  mutate(group = ifelse(record_id %in% groupA,"A","B"))



df_photo <- bind_rows(df_descr %>%
                        select(id_cc, total_photographs_sent_26,photo_compliance_week_1) %>%
                        slice(1:45) %>% 
                        mutate(week = 1) %>%
                        setNames(c("record_id","total photographs","photo compliance","week")),
                      df_descr %>%
                        select(id_cc, total_photographs_sent_49,photo_compliance_week_2) %>%
                        slice(1:45) %>%
                        mutate(week = 2) %>%
                        setNames(c("record_id","total photographs","photo compliance","week"))
                      ) %>%
  mutate(group=as.factor(ifelse(record_id %in% groupA, "A","B"))) %>%
  mutate(phase=case_when(group=="A" & week==1~"Morning", 
                         group=="A" & week==2~"Evening",
                         group=="B" & week==1~"Evening",
                         group=="B" & week==2~"Morning")) %>%
  mutate(record_id=as.factor(record_id))
  


df_age <- df_descr %>%
  select(id_cc,elikia_age) %>%
  slice(1:45) %>%
  setNames(c("record_id","age"))

df_info <- df %>% select("record_id","gender","city","group") %>% # filtering out the personal info
  mutate(city=ifelse(city==2,"Limassol","Paphos"),
         gender=ifelse(gender==2,"Female","Male")) %>%
  left_join(df_age) %>%
  mutate(record_id= as.factor(record_id))

df_photo <- df_info %>%
  left_join(df_photo)

df_anthrop <- df %>%
  select(
    record_id, height, weight, bmi, waist,
    weight_2, bmi_2, waist_2,
    weight_3, bmi3, waist_3, group
  ) %>%
  mutate(height = as.numeric(ifelse(height < 100, height * 100, height))) %>%
  mutate(record_id = as.factor(record_id)) %>%
  left_join(df_info)

```

```{r splitting the daily questionnaires, include=FALSE}


df_day1 <- df %>%
  select(record_id, ccaa1_timestamp:ccaa1_complete) %>%
  mutate(day=1) %>%
  select(-all_of(c(d1_extra,"urine_sample")))

df_day2 <- df %>%
  select(record_id, cedd2_timestamp:cedd2_complete) %>%
  mutate(day=2) %>%
  setNames(names(df_day1))

df_day3 <- df %>%
  select(record_id, eccb3_timestamp:eccb3_complete) %>%
  mutate(day=3) %>%
  select(-"urine_sample_d3") %>%
  setNames(names(df_day1))

df_day4 <- df %>%
  select(record_id, affa4_timestamp:affa4_complete) %>%
  mutate(day=4) %>%
  setNames(names(df_day1))

df_day5 <- df %>%
  select(record_id, edab5_timestamp:edab5_complete) %>%
  mutate(day=5) %>%
  select(-"urine_sample_d5") %>%
  setNames(names(df_day1))

df_day6 <- df %>%
  select(record_id, cafa6_timestamp:cafa6_complete) %>%
  mutate(day=6) %>%
  setNames(names(df_day1))

df_day7 <- df %>%
  select(record_id, fefc7_timestamp:fefc7_complete) %>%
  mutate(day=7) %>%
  select(-"urine_sample_d7") %>% 
  setNames(names(df_day1))

df_day8 <- df %>%
  select(record_id, ccaa1_b2a5_timestamp:ccaa1_b2a5_complete) %>%
  mutate(day=8) %>%
  select(-all_of(c(d8_extra,"urine_sample_d8"))) %>%
  setNames(names(df_day1))

df_day9 <- df %>%
  select(record_id, cedd2_2e8d_timestamp:cedd2_2e8d_complete) %>%
  mutate(day=9) %>%
  setNames(names(df_day1))

df_day10 <- df %>%
  select(record_id, eccb3_c150_timestamp:eccb3_c150_complete) %>%
  mutate(day=10) %>%
  select(-"urine_sample_d10") %>%
  setNames(names(df_day1))

df_day11 <- df %>%
  select(record_id, affa4_a7f6_timestamp:affa4_a7f6_complete) %>%
  mutate(day=11) %>%
  setNames(names(df_day1))

df_day12 <- df %>%
  select(record_id, edab5_b029_timestamp:edab5_b029_complete) %>%
  mutate(day=12) %>%
  select(-"urine_sample_d12") %>%
  setNames(names(df_day1))

df_day13 <- df %>% 
  select(record_id, cafa6_3763_timestamp:cafa6_3763_complete) %>%
  mutate(day=13) %>%
  setNames(names(df_day1))
  

df_day14 <- df %>% 
  select(record_id, fefc7_c416_timestamp:fefc7_c416_complete) %>%
  mutate(day=14) %>%
  select(-"urine_sample_d14") %>%
  setNames(names(df_day1))

```


```{r creating the urine sample dataset, include=FALSE}
#urine sample data
df_urine <- df %>%
  select(
    urine_sample, urine_sample_d3, urine_sample_d5, urine_sample_d7,
    urine_sample_d8, urine_sample_d10, urine_sample_d12, urine_sample_d14
  ) %>%
  setNames(c("day_1", "day_3", "day_5", "day_7", "day_8", "day_10", "day_12", "day_14")) %>%
  mutate(across(everything(), ~ replace(., . == 1, "Yes"))) %>%
  mutate(across(everything(), ~ replace(., . == 2, "No")))
```

```{r , include=FALSE}
#extra information for day1
d1_extra <- c("onset_date","food_supplements","smoking_general",
              "smoking_products","cigarettes_mean","supplements_diary")

#extra information for day8
d8_extra <- c("onset_date_d8","food_supplements_d8","smoking_general_d8",
              "smoking_products_d8","cigarettes_mean_d8","supplements_diary_d8")

df_extra <- bind_rows(df %>%
                        select(all_of(d1_extra)) %>%
                        mutate(day=1),
                      df %>%
                        select(all_of(d8_extra)) %>%
                        mutate(day=8) %>%
                        setNames(c(d1_extra,"day")))


dataset <- bind_rows(
  df_day1,df_day2,df_day3,df_day4,df_day5, df_day6, df_day7,
  df_day8, df_day9, df_day10, df_day11, df_day12, df_day13, df_day14
)

rm(df_day1,df_day2,df_day3,df_day4,df_day5, df_day6, df_day7,
  df_day8, df_day9, df_day10, df_day11, df_day12, df_day13, df_day14)
```

```{r, include=FALSE}
#creating a function to change the index of fruits and vegetables
fruit <- function(x){
  x=case_when(x==1 ~ 1,
              x==2 ~ 0.75,
              x==3 ~ 0.5,
              x==4 ~ 0.25,
              x==5 ~ 0,
              x==6 ~ -1)
  return(x)
}

#change the index of fruits and vegetables using the custom function
dataset <- dataset %>% mutate(orange=fruit(orange),
                              banana=fruit(banana),
                              apple=fruit(apple),
                              pear=fruit(pear),
                              tomatoes=fruit(tomatoes),
                              lettuce=fruit(lettuce),
                              cucumber=fruit(cucumber))

#we code the 6 as -1. -1 is the different proportion of consumption where participants asked to filled the exact amount. So we combining the two columns to one.
dataset[dataset$orange==-1,"orange"] <- dataset[dataset$orange==-1,"other_orange"]
dataset[dataset$apple==-1,"apple"] <- dataset[dataset$apple==-1,"other_apple"]
dataset[dataset$cucumber==-1,"cucumber"] <- dataset[dataset$cucumber==-1,"other_cucumber"]/2

#calculating the fruit, veg and total consumption and the prop of consumption
dataset <- dataset %>% mutate(fruit=orange+banana+apple+pear,
                              veg=tomatoes+lettuce+cucumber,
                              fruit_veg=fruit+veg,
                              prop_fv=fruit_veg/7,
                              prop_f=fruit/4,
                              prop_v=veg/3)

#compliance metrics.
dataset <- dataset %>% mutate(time_comp=orange_time___2+orange_time___5 + banana_time___2 +
                                banana_time___5 + apple_time___2 + apple_time___5 +
                                pear_time___2 + pear_time___5 + cucumber_time___2+
                                cucumber_time___5 + tomatoe_time___2 + tomatoe_time___5+
                                lettuce_time___2 + lettuce_time___5,
                              time_comp_prop=time_comp/7,
                              time912= orange_time___2 + banana_time___2 + apple_time___2 + 
                                pear_time___2 + cucumber_time___2 + tomatoe_time___2 + 
                                lettuce_time___2,
                              time710=orange_time___5 + banana_time___5 + apple_time___5 + 
                                pear_time___5 + cucumber_time___5 + tomatoe_time___5 + 
                                lettuce_time___5)
```

```{r, include=FALSE}
dfr <- dataset %>%
  mutate(
    week = ifelse(day %in% 1:7, "1", "2"), # adding the week
    record_id = as.factor(record_id), # adding the record id
    group = as.factor(ifelse(record_id %in% groupA, "A", "B")), # adding group
    phase = case_when(
      group == "A" & week == "1" ~ "Morning", # adding phase
      group == "A" & week == "2" ~ "Evening",
      group == "B" & week == "1" ~ "Evening",
      group == "B" & week == "2" ~ "Morning"
    )
  ) %>%
  select(
    record_id, group, day, week, fruit, veg, fruit_veg, prop_fv, prop_f, prop_v,
    time_comp, time_comp_prop, time912, time710, phase
  )

```


```{r, include=FALSE}
# Typically the participants need to consume 7 fruits/veg for 7 days, so in total 49. Participants with consumption below 80%  need to be excluded. 49*0.8=39.2. So participants below 39 in both weeks need to be excluded.

#participants with consumption below 39 in first week 
fir <- dfr %>%
  filter(week==1) %>%
  group_by(record_id) %>% 
  summarise(sum=sum(fruit_veg)) %>%
  ungroup() %>%
  filter(sum<39)

#participants with consumption below 39 in second week
sec <- dfr %>%
  filter(week==2) %>%
  group_by(record_id) %>% 
  summarise(sum=sum(fruit_veg)) %>%
  ungroup() %>%
  filter(sum<39)

#participants with consumption below 39 in both weeks.
firsec <- intersect(fir$record_id,sec$record_id) # these participants will be excluded
```

### id with consumption below 39 for first week

`r kable(print(fir), "pipe")`

### id with consumption below 39 for second week

`r kable(print(sec), "pipe")`

### id with consumption below 39 in both weeks

`r kable(print(firsec),"pipe")`

```{r, include=FALSE}
dfr <- dfr %>%
  filter(!(record_id %in% firsec)) %>%
  mutate(record_id=factor(record_id))

df_info <- df_info %>%
  filter(!(record_id %in% firsec)) %>%
  mutate(record_id=factor(record_id))

df_anthrop <- df_anthrop %>%
  filter(!(record_id %in% firsec)) %>%
  mutate(record_id=factor(record_id))

df_photo <- df_photo %>%
  filter(!(record_id %in% firsec)) %>%
  mutate(record_id=factor(record_id))

dataset <- dataset %>%
  filter(!(record_id %in% firsec)) %>%
  mutate(record_id=factor(record_id))

df <- df %>%
  filter(!(record_id %in% firsec)) %>%
  mutate(record_id=factor(record_id))

rm(df_urine, df_extra,fir,sec,firsec)
```




`r kable(print(CreateTableOne(c("gender","city","age"), data=df_info, strata=c("group"),,addOverall=T)), "pipe")`

`r kable(print(CreateTableOne(c("age"), data=df_info, strata=c("gender"), ,addOverall=T)), "pipe")`

`r kable(print(CreateTableOne(c("height","weight","weight_2","weight_3","waist","waist_2","waist_3","bmi","bmi_2", "bmi3"), data=df_anthrop, strata=c("group"))), "pipe")`

`r kable(print(CreateTableOne(c("height","weight","weight_2","weight_3","waist","waist_2","waist_3","bmi","bmi_2", "bmi3"), data=df_anthrop, strata=c("gender"))), "pipe")`

### Photo compliance

`r kable(print(CreateTableOne(c("total photographs","photo compliance"), data=df_photo, strata=c("group")),nonnormal = c("total photographs")), "pipe")`

`r kable(print(CreateTableOne(c("total photographs","photo compliance"), data=df_photo, strata=c("gender")),nonnormal = c("total photographs")), "pipe")`

`r kable(print(CreateTableOne(c("total photographs","photo compliance"), data=df_photo, strata=c("phase"), testNonNormal = wilcox.test , argsNonNormal = list(paired = FALSE), testNormal=t.test, argsNormal=list(paired = FALSE)),nonnormal = c("total photographs")), "pipe")`

`r kable(print(CreateTableOne(c("total photographs","photo compliance"), data=df_photo, strata=c("week"), testNonNormal = wilcox.test , argsNonNormal = list(paired = TRUE), testNormal=t.test, argsNormal=list(paired=TRUE)),nonnormal = c("total photographs")), "pipe")`



### Compliance tables

#### Compliance table for participant who suppose to eat the veg/fruits 7-10 pm
`r kable(print(table(dataset$phase, dataset$time710)),  "pipe")`

#### Compliance table for participant who suppose to eat the veg/fruits 9-12 am

`r kable(print(table(dataset$phase, dataset$time912)),  "pipe")`

### Summary of fruit, veg, sum and proportion per group type and week

`r kable(print(CreateTableOne(c("fruit","veg","fruit_veg","prop_fv","prop_f","prop_v"), data=dfr, strata=c("week") ,addOverall=T,testNonNormal = wilcox.test , argsNonNormal = list(paired = TRUE), testNormal=t.test, argsNormal=list(paired=TRUE)),nonnormal = c("fruit","veg","fruit_veg")),  "pipe")`

`r kable(print(CreateTableOne(c("fruit","veg","fruit_veg","prop_fv","prop_f","prop_v"), data=dfr, strata=c("group") ,addOverall=T),nonnormal = c("fruit","veg","fruit_veg")), "pipe")`

`r kable(print(CreateTableOne(c("fruit","veg","fruit_veg","prop_fv","prop_f","prop_v"), data=dfr, strata=c("phase") ,addOverall=T,testNonNormal = wilcox.test , argsNonNormal = list(paired = TRUE), testNormal=t.test, argsNormal=list(paired=TRUE)),nonnormal = c("fruit","veg","fruit_veg")), "pipe")`

### Summary of time compliance per Group type and week.

`r kable(print(CreateTableOne(c("time_comp", "time_comp_prop"), data=dfr, strata=c("group") ,addOverall=T),nonnormal = c("time_comp")), "pipe")`

`r kable(print(CreateTableOne(c("time_comp", "time_comp_prop"), data=dfr, strata=c("week") ,addOverall=T,testNonNormal = wilcox.test , argsNonNormal = list(paired = TRUE), testNormal=t.test, argsNormal=list(paired=TRUE)),nonnormal = c("time_comp")), "pipe")`

`r kable(print(CreateTableOne(c("time_comp", "time_comp_prop"), data=dfr, strata=c("phase") ,addOverall=T,testNonNormal = wilcox.test , argsNonNormal = list(paired = TRUE), testNormal=t.test, argsNormal=list(paired=TRUE)),nonnormal = c("time_comp")), "pipe")`

```{r ,include=FALSE, message=FALSE, warning=FALSE}
df_cal <- read_excel("./data_files/Fruits-Vegetables-kcal.xlsx",skip = 1) %>%
  setNames(c("record_id","group","fruits_per_day","veg_per_day","fruits_per_week",
                    "veg_per_week","total_daily_fv","kcal_bl","kcal_1st","kcal_2nd",
                   "water_bl","water_1st","water_2nd","rem")) %>%
  select(-rem) %>%
  filter(!record_id %in% c(excluded_patients,firsec)) %>%
  mutate(record_id = as.factor(record_id),
         kcal_1st = as.numeric(kcal_1st),
         water_bl = as.numeric(water_bl))

fv_7 <- dfr %>%
  filter(day==7) %>%
  select(record_id,fruit_veg) %>% 
  filter(!(record_id %in% c(excluded_patients,firsec))) %>%
  mutate(kcal_7=round((fruit_veg/7)*443)) %>%
  rename(fruit_veg_7=fruit_veg)

fv_14 <- dfr %>%
  filter(day==14) %>%
  select(record_id,fruit_veg) %>% 
  filter(!(record_id %in% c(excluded_patients,firsec))) %>%
  mutate(kcal_14=round((fruit_veg/7)*443)) %>%
  rename(fruit_veg_14=fruit_veg)

df_cal2 <- df_cal %>%
  left_join(fv_7) %>%
  left_join(fv_14) %>%
  filter(!is.na(kcal_1st) & !is.na(kcal_2nd)) %>%
  mutate(fv_prop_cal_1st=round((kcal_7/kcal_1st)*100)) %>%
  mutate(fv_prop_cal_2nd=round((kcal_14/kcal_2nd)*100))


df_cal3 <- df_cal2 %>%
  select(record_id,group,kcal_bl,kcal_1st,kcal_2nd) %>%
  reshape2::melt() %>%
  mutate(week=case_when(variable=="kcal_bl" ~1,
                        variable=="kcal_1st" ~2,
                        variable=="kcal_2nd"~3)) %>%
  mutate(phase=case_when(group=="A" & week=="2"~"Morning",
                         group=="A" & week=="3"~"Evening",
                         group=="B" & week=="2"~"Evening",
                         group=="B" & week=="3"~"Morning",
                         group=="A" & week=="1"~"Baseline",
                         group=="B" & week=="1"~"Baseline")) %>%
  rename(kcal=value)

df_cal4 <- df_cal2 %>%
  select(record_id,group,fv_prop_cal_1st, fv_prop_cal_2nd) %>%
  reshape2::melt() %>%
  mutate(week=case_when(variable=="fv_prop_cal_1st" ~1,
                        variable=="fv_prop_cal_2nd" ~2)) %>%
  mutate(phase=case_when(group=="A" & week=="2"~"Morning",
                         group=="A" & week=="3"~"Evening",
                         group=="B" & week=="2"~"Evening",
                         group=="B" & week=="3"~"Morning")) %>%
  rename(fruit_veg_prop=value)

kcal_diff <- df_cal2 %>%
  select(record_id,group,kcal_bl,kcal_1st,kcal_2nd) %>%
  mutate(kcal1b=kcal_1st-kcal_bl, 
         kcal21=kcal_2nd-kcal_1st,
         kcal2b=kcal_2nd-kcal_bl) %>%
  select(record_id,kcal1b,kcal21,kcal2b,group) %>%
  reshape2::melt() %>%
  mutate(phase=case_when(group=="A" & variable=="kcal1b"~"Morning",
                         group=="B" & variable=="kcal1b"~"Evening",
                         group=="A" & variable=="kcal21"~"Evening",
                         group=="B" & variable=="kcal21"~"Morning",
                         group=="A" & variable=="kcal2b"~"Total",
                         group=="B" & variable=="kcal2b"~"Total")) %>%
  dplyr::rename(kcal_difference=value)

kcal_diff2 <- kcal_diff %>%
  filter(phase!="Total")

df_cal <- df_cal %>%
  mutate(total_weekly_fv=fruits_per_week+veg_per_week)


cal_descr <- df_cal %>%
  select(kcal_bl,kcal_1st,kcal_2nd) %>%
  psych::describe() %>%
  select(n, mean, sd, median, se, min, max, range)

```

### FFQ descriptives stratified by Group

`r kable(print(CreateTableOne(c("fruits_per_day","veg_per_day","fruits_per_week", "veg_per_week","total_daily_fv","total_weekly_fv","kcal_bl","kcal_1st","kcal_2nd"), data=df_cal, strata=c("group"),,addOverall=T)), "pipe")`


### Calories

#### Calories descriptives

`r kable(print(cal_descr), "pipe")`

#### Calories per phase

`r kable(print(CreateTableOne(c("kcal"), data=df_cal3, strata=c("phase") ,addOverall=T)), "pipe")`

#### Calories per group

`r kable(print(CreateTableOne(c("kcal"), data=df_cal3, strata=c("group") ,addOverall=T)), "pipe")`

#### Calories percentage from fruits and vegetables per phase

`r kable(print(CreateTableOne(c("fruit_veg_prop"), data=df_cal4, strata=c("phase") ,addOverall=T)), "pipe")`

#### Calories percentage from fruits and vegetables per group

`r kable(print(CreateTableOne(c("fruit_veg_prop"), data=df_cal4, strata=c("group") ,addOverall=T)), "pipe")`

#### Calories difference per phase

`r kable(print(CreateTableOne(c("kcal_difference"), data=kcal_diff2, strata=c("phase"), testNormal=t.test, argsNormal=list(paired=TRUE))), "pipe")`

