---
title: "Circa chem"
author: "Stavros Oikonomou"
date: "11/8/2021"
output:
  html_document: default
  word_document: default
---

```{r setup II, include=FALSE, eval=TRUE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, include = FALSE)
```

```{r, warning=FALSE, message=FALSE, include=FALSE}
rm(list = ls(all = TRUE))
ipak <- function(pkg) {
  new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
  if (length(new.pkg)) {
    install.packages(new.pkg, dependencies = TRUE)
  }
  sapply(pkg, require, character.only = TRUE)
}

# usage
packages <- c(
  "readxl", "dplyr", "stringr", "NADA", "lme4", "sjPlot",
  "tableone", "knitr", "ggplot2", "reshape2"
)

ipak(packages)
```


```{r, include=FALSE}
# loading the datasets
pba_cn <- read_excel("./data_files/3PBA-6CN Creatinine CIRCACHEM Final.xlsx",
  sheet = "6CN-3PBA", skip = 1,
  col_names = c("Code", "CN", "PBA", "X4", "X5", "X6", "X7")
)

df_crea <- read_excel("./data_files/3PBA-6CN Creatinine CIRCACHEM Final.xlsx",
  sheet = "Creatinine", col_names = c("Code", "Creatinine"), skip = 1
)

df_hne <- read_excel("./data_files/4HNE CIRCACHEM RESULTS_v2.xlsx",
  sheet = "final3", col_names = c("Code", "hne"), skip = 1
)

df_descr <- read_excel("./data_files/CIRCA CHEM_Participants Registry_29.12.20.xlsx")
```

```{r,include=FALSE}
# setting the cn/pb lod and loq
cn_lod <- 0.075
cn_loq <- 0.226
pb_lod <- 0.049
pb_loq <- 0.146
```

```{r, include=FALSE}
# creating the info from Code feature
info_pest <- str_split_fixed(pba_cn$Code, "-", 4) %>%
  data.frame() %>%
  select(-X1) %>%
  rename(record_id = X2, group = X3, sample_number = X4) %>%
  mutate(
    record_id = as.factor(record_id),
    group = as.factor(group),
    sample_number = as.factor(sample_number)
  )


# bind back to original the splitted information
df_pest <- cbind.data.frame(info_pest, CN = pba_cn$CN, PBA = pba_cn$PBA)

# creating variables for descriptive
df_pest <- df_pest %>% mutate(
  cnlod = as.factor(ifelse(CN < cn_lod, "Below LOD", "Above LOD")),
  pbalod = as.factor(ifelse(PBA < pb_lod, "Below LOD", "Above LOD")),
  cnlod_TF = ifelse(CN < cn_lod, T, F),
  cnlodloq = cut(CN,
    breaks = c(-Inf, cn_lod, cn_loq, Inf),
    labels = c("ND", "NQ", "D&Q")
  ),
  pblodloq = cut(PBA,
    breaks = c(-Inf, pb_lod, pb_loq, Inf),
    labels = c("ND", "NQ", "D&Q")
  )
)

prop.table(table(df_pest$cnlod))
prop.table(table(df_pest$cnlodloq))

prop.table(table(df_pest$pbalod))
prop.table(table(df_pest$pblodloq))

# applying ros function to deal with values below LOD. We going to use ros function because with have a big percentage of values below lod.
cnf <- ros(df_pest$CN, df_pest$cnlod_TF, forwardT = "log", reverseT = "exp", na.action = NULL) %>%
  as.data.frame()

# bind back the new values
df_pest2 <- merge(df_pest, cnf, by.x = "CN", by.y = "obs") %>%
  select(record_id, group, sample_number, PBA, modeled) %>%
  dplyr::rename(CN = modeled) %>%
  arrange(record_id, group, sample_number)

# check PBA and CN for normality
hist(df_pest2$PBA)
hist(df_pest2$CN)

rm(pba_cn, info_pest, cnf)
```

`r kable(prop.table(table(df_pest$cnlod)),caption="6-CN Above and Below LOD percentage")`

`r kable(prop.table(table(df_pest$cnlodloq)),caption="6-CN Above and Below LOD/LOQ/Above LOQ percentage")`

`r kable(prop.table(table(df_pest$pbalod)),caption="3-PBA Above and Below LOD percentage")`

`r kable(prop.table(table(df_pest$pblodloq)),caption="3-PBA Above and Below LOD/LOQ/Above LOQ percentage")`


```{r, include=FALSE}
# creating the info from Code feature
cr_info <- str_split_fixed(df_crea$Code, "-", 4) %>%
  data.frame() %>%
  select(-X1) %>%
  rename(record_id = X2, group = X3, sample_number = X4) %>%
  mutate(
    record_id = as.factor(record_id),
    group = as.factor(group),
    sample_number = as.factor(sample_number)
  ) %>%
  # we need to change the one digit entries for 1 to 01 in order to match with the PBA/CN dataset
  mutate(record_id = as.factor(ifelse(str_length(record_id) == 1, paste0("0", record_id),
    as.character(record_id)
  )))

# bind back the data
df_cr <- cbind(cr_info, creatinine = df_crea$Creatinine) %>%
  arrange(record_id, group, sample_number)

# checking the normality of creatinine
hist(df_cr$crea)

# counting the number of entries per participant
removing_part <- df_cr %>%
  group_by(record_id) %>%
  summarise(n = n()) %>%
  ungroup() %>%
  filter(n < 8) %>% # participants with not 8 entries need to be excluded
  pull(record_id)


rm(cr_info, df_crea, df_pest)
```

```{r, include=FALSE, warning=FALSE, message=FALSE}
# creating the info for hne data
hne_info <- str_split_fixed(df_hne$Code, "-", 4) %>%
  data.frame() %>%
  select(-X1) %>%
  rename(record_id = X2, group = X3, sample_number = X4) %>%
  mutate(
    record_id = as.factor(record_id),
    group = as.factor(group),
    sample_number = as.factor(sample_number)
  ) %>%
  # we need to change the one digit entries for 1 to 01 in order to match with the PBA/CN dataset
  mutate(record_id = as.factor(ifelse(str_length(record_id) == 1, paste0("0", record_id),
    as.character(record_id)
  )))



df_hne2 <- cbind(hne_info, hne = df_hne$hne) %>%
  arrange(record_id, group, sample_number) %>%
  mutate(hne = as.numeric(hne))

hist(df_hne2$hne)

rm(hne_info, df_hne, cn_lod, cn_loq, pb_lod, pb_loq)
```


```{r, include=FALSE}
# age and sex data
df_age_sex <- df_descr[1:45, c(1, 6, 7)] %>%
  setNames(., c("record_id", "sex", "age")) %>%
  mutate(
    record_id = as.factor(record_id),
    sex = as.factor(ifelse(sex == "F", "Female", "Male"))
  ) %>%
  mutate(record_id = as.factor(ifelse(str_length(record_id) == 1, paste0("0", record_id),
    as.character(record_id)
  )))

rm(df_descr)
```

```{r, include=FALSE}
# combine PBA/CN/HNE and creatinine dataset.
df <- merge(df_pest2, df_cr, all = T, by = c("record_id", "group", "sample_number")) %>%
  merge(., df_hne2, all = T, by = c("record_id", "group", "sample_number")) %>%
  filter(!(record_id %in% removing_part)) %>% # removing non complete data participant
  filter(!(record_id %in% c("43", "32"))) %>% # removing participants from compliance filter in previous script
  mutate(record_id = factor(record_id)) %>%
  mutate(
    cn_adj = CN / creatinine, # adjust measurements with creatinine
    pb_adj = PBA / creatinine,
    hne_adj = hne / creatinine
  ) %>%
  # creating the phase info
  mutate(phase = case_when(
    group == "A" & sample_number %in% c("1", "2", "3", "4") ~ "Morning",
    group == "A" & sample_number %in% c("5", "6", "7", "8") ~ "Evening",
    group == "B" & sample_number %in% c("1", "2", "3", "4") ~ "Evening",
    group == "B" & sample_number %in% c("5", "6", "7", "8") ~ "Morning"
  ))

df_nobl <- df %>% filter(!(sample_number %in% c("1", "5"))) # non baseline data

df_bl <- df %>% filter(sample_number %in% c("1", "5")) %>% # baseline data
  select(-sample_number)

# combining baseline and treatment data
df_final <- merge(df_nobl, df_bl,
  by = c("record_id", "group", "phase"), all.x = T,
  suffixes = c("_TR", "_BL")
) %>%
  mutate(sample_number_order = case_when(
    sample_number %in% c("2", "6") ~ "1",
    sample_number %in% c("3", "7") ~ "2",
    sample_number %in% c("4", "8") ~ "3"
  )) %>%
  # log and scale the adjusted cn,pba and hne measurements
  mutate(
    cn_adj_TR_log_S = scale(log(cn_adj_TR), scale = FALSE),
    cn_adj_BL_log_S = scale(log(cn_adj_BL), scale = FALSE),
    pb_adj_TR_log_S = scale(log(pb_adj_TR), scale = FALSE),
    pb_adj_BL_log_S = scale(log(pb_adj_BL), scale = FALSE),
    hne_adj_TR_log_S = scale(log(hne_adj_TR), scale = FALSE),
    hne_adj_BL_log_S = scale(log(hne_adj_BL), scale = FALSE)
  ) %>%
  # join with age and sex data.
  left_join(df_age_sex)

rm(df_bl, df_nobl)

df2 <- na.omit(df)
df2 <- df2[-98, ]
```

`r kable(print(CreateTableOne(c("PBA","CN","creatinine","hne"), data=df2,  strata=c("phase"), addOverall=T, testNonNormal = wilcox.test , argsNonNormal = list(paired = TRUE)), nonnormal =c("PBA","CN","creatinine","hne")), "pipe")`

```{r plots, echo=FALSE, warning=FALSE, message=FALSE}
ggplot(df, aes(x = phase, y = CN, fill = phase)) +
  geom_boxplot() +
  xlab("Phase") +
  ylab("CN") +
  theme_minimal() +
  ggtitle("6-CN boxplot for morning and evening")

ggplot(df, aes(x = phase, y = PBA, fill = phase)) +
  geom_boxplot() +
  xlab("Phase") +
  ylab("PBA") +
  theme_minimal() +
  ggtitle("3-PBA boxplot for morning and evening")

ggplot(df, aes(x = phase, y = creatinine, fill = phase)) +
  geom_boxplot() +
  xlab("Phase") +
  ylab("Creatinine") +
  theme_minimal() +
  ggtitle("Creatinine boxplot for morning and evening")

ggplot(df, aes(x = phase, y = hne, fill = phase)) +
  geom_boxplot() +
  xlab("Phase") +
  ylab("4-HNE") +
  theme_minimal() +
  ggtitle("4-HNE boxplot for morning and evening")
```

```{r normality check, echo=FALSE}
# raw data qqplot
qqnorm(df$CN, main = "6-CN QQ plot")
qqnorm(df$PBA, main = "3-PBA QQ plot")
qqnorm(df$hne, main = "4-HNE QQ plot")

# adj by creatinine qqplot
qqnorm(df$cn_adj, main = "6-CN adjusted by creatinine QQ plot")
qqnorm(df$pb_adj, main = "3-PBA adjusted by creatinine QQ plot")
qqnorm(df$hne_adj, main = "4-HNE adjusted by creatinine QQ plot")

# log transformed for adjusted data
df <- df %>% mutate(
  cn_logadj = log(cn_adj),
  pb_logadj = log(pb_adj),
  hne_logadj = log(hne_adj)
)

# qqplot for log transformed adjusted data.
qqnorm(df$cn_logadj, main = "6-CN adjusted by creatinine and log transformed QQ plot")
qqnorm(df$pb_logadj, main = "3-PBA adjusted by creatinine and log transformed QQ plot")
qqnorm(df$hne_logadj, main = "4-HNE adjusted by creatinine and log transformed QQ plot")
```

```{r plots2, echo=FALSE, message=FALSE, warning=FALSE}
# CN plos
ggplot(df_final, aes(x = record_id, y = cn_adj_TR_log_S, fill = phase)) +
  geom_boxplot() +
  xlab("Participants") +
  ylab("6-CN adjusted,log transformed and centered") +
  theme_minimal() +
  geom_point(aes(x = record_id, y = cn_adj_BL_log_S, colour = phase), size = 2)

ggplot(df_final, aes(x = record_id, y = cn_adj_TR_log_S, fill = phase)) +
  geom_boxplot() +
  xlab("Participants") +
  ylab("6-CN adjusted,log transformed and centered") +
  theme_minimal() +
  facet_wrap(~phase)

ggplot(df_final, aes(x = phase, y = cn_adj_TR_log_S, fill = phase)) +
  geom_boxplot() +
  theme_minimal()
ggplot(df_final, aes(x = phase, y = cn_adj_BL_log_S, fill = phase)) +
  geom_boxplot() +
  theme_minimal()


# pb plots
ggplot(df_final, aes(x = record_id, y = pb_adj_TR_log_S, fill = phase)) +
  geom_boxplot() +
  xlab("Participants") +
  ylab("3-PBA adjusted,log transformed and centered") +
  theme_minimal() +
  geom_point(aes(x = record_id, y = pb_adj_BL_log_S, colour = phase), size = 2)

ggplot(df_final, aes(x = record_id, y = pb_adj_TR_log_S, fill = phase)) +
  geom_boxplot() +
  xlab("Participants") +
  ylab("3-PBA adjusted,log transformed and centered") +
  theme_minimal() +
  facet_wrap(~phase)

ggplot(df_final, aes(x = phase, y = pb_adj_TR_log_S, fill = phase)) +
  geom_boxplot() +
  theme_minimal()
ggplot(df_final, aes(x = phase, y = pb_adj_BL_log_S, fill = phase)) +
  geom_boxplot() +
  theme_minimal()

# hne
ggplot(df_final, aes(x = record_id, y = hne_adj_TR_log_S, fill = phase)) +
  geom_boxplot() +
  xlab("Participants") +
  ylab("4-HNE adjusted,log transformed and centered") +
  theme_minimal() +
  geom_point(aes(x = record_id, y = hne_adj_BL_log_S, colour = phase), size = 2)

ggplot(df_final, aes(x = record_id, y = hne_adj_TR_log_S, fill = phase)) +
  geom_boxplot() +
  xlab("Participants") +
  ylab("4-HNE adjusted,log transformed and centered") +
  theme_minimal() +
  facet_wrap(~phase)

ggplot(df_final, aes(x = phase, y = hne_adj_TR_log_S, fill = phase)) +
  geom_boxplot() +
  theme_minimal()
ggplot(df_final, aes(x = phase, y = hne_adj_BL_log_S, fill = phase)) +
  geom_boxplot() +
  theme_minimal()
```

```{r, include=FALSE}
df_final <- df_final %>%
  mutate(phase = as.factor(phase)) %>%
  mutate_if(grepl("_log_S", names(df_final)), as.numeric) %>%
  mutate(days = if_else(sample_number_order=="1",2,
                        if_else(sample_number_order=="2",4,7))) %>%
  mutate(days = case_when(record_id == "28" & sample_number == "8" ~ 8,
                          record_id == "02" & sample_number == "6" ~ 1,
                          record_id == "02" & sample_number == "7" ~ 5,
                          TRUE ~ days)) 



df_final2 <- na.omit(df_final)
df_final2 <- df_final2[-73, ]
```

`r kable(print(CreateTableOne(c("CN_BL","CN_TR","PBA_BL","PBA_TR","hne_BL","hne_TR","cn_adj_BL","cn_adj_TR","pb_adj_BL", "pb_adj_TR", "hne_adj_BL","hne_adj_TR"), data=df_final2, strata=c("phase"),addOverall=T ,testNonNormal = wilcox.test, argsNonNormal = list(paired = TRUE) ),nonnormal =c("CN_BL","CN_TR","PBA_BL","PBA_TR","hne_BL","hne_TR","cn_adj_BL","cn_adj_TR","pb_adj_BL", "pb_adj_TR","hne_adj_BL","hne_adj_TR")), "pipe")`

```{r quantiles, include=FALSE}


cn_qo <- round(quantile(df$CN, c(0, .05, .25, .5, .75, .95, 1)), digits = 2)
pb_qo <- round(quantile(df$PBA, c(0, .05, .25, .5, .75, .95, 1)), digits = 2)
cr_qo <- round(quantile(df$creatinine, c(0, .05, .25, .5, .75, .95, 1)), digits = 2)

hne_qo <- round(quantile(na.omit(df$hne), c(0, .05, .25, .5, .75, .95, 1)), digits = 2)

cnbl_qo <- round(quantile(df_final$CN_BL, c(0, .05, .25, .5, .75, .95, 1)), digits = 2)
pbbl_qo <- round(quantile(df_final$PBA_BL, c(0, .05, .25, .5, .75, .95, 1)), digits = 2)
hnebl_qo <- round(quantile(df_final$hne_BL, c(0, .05, .25, .5, .75, .95, 1)), digits = 2)

cntr_qo <- round(quantile(df_final$CN_TR, c(0, .05, .25, .5, .75, .95, 1)), digits = 2)
pbtr_qo <- round(quantile(df_final$PBA_TR, c(0, .05, .25, .5, .75, .95, 1)), digits = 2)
hnetr_qo <- round(quantile(na.omit(df_final$hne_TR), c(0, .05, .25, .5, .75, .95, 1)), digits = 2)

cnadjbl_qo <- round(quantile(df_final$cn_adj_BL, c(0, .05, .25, .5, .75, .95, 1)), digits = 2)
pbadjbl_qo <- round(quantile(df_final$pb_adj_BL, c(0, .05, .25, .5, .75, .95, 1)), digits = 2)
hneadjbl_qo <- round(quantile(df_final$hne_adj_BL, c(0, .05, .25, .5, .75, .95, 1)), digits = 2)

cnadjtr_qo <- round(quantile(df_final$cn_adj_TR, c(0, .05, .25, .5, .75, .95, 1)), digits = 2)
pbadjtr_qo <- round(quantile(df_final$pb_adj_TR, c(0, .05, .25, .5, .75, .95, 1)), digits = 2)
hneadjtr_qo <- round(quantile(na.omit(df_final$hne_adj_TR), c(0, .05, .25, .5, .75, .95, 1)), digits = 2)

cnadjbls_qo <- round(quantile(df_final$cn_adj_BL_log_S, c(0, .05, .25, .5, .75, .95, 1)), digits = 2)
pbadjbls_qo <- round(quantile(df_final$pb_adj_BL_log_S, c(0, .05, .25, .5, .75, .95, 1)), digits = 2)
hneadjbls_qo <- round(quantile(df_final$hne_adj_BL_log_S, c(0, .05, .25, .5, .75, .95, 1)), digits = 2)

cnadjtrs_qo <- round(quantile(df_final$cn_adj_TR_log_S, c(0, .05, .25, .5, .75, .95, 1)), digits = 2)
pbadjtrs_qo <- round(quantile(df_final$pb_adj_TR_log_S, c(0, .05, .25, .5, .75, .95, 1)), digits = 2)
hneadjtrs_qo <- round(quantile(na.omit(df_final$hne_adj_TR_log_S), c(0, .05, .25, .5, .75, .95, 1)), digits = 2)

quantil <- rbind(
  cn_qo, pb_qo, cr_qo, hne_qo, cnbl_qo, pbbl_qo, hnebl_qo, cntr_qo, pbtr_qo, hnetr_qo,
  cnadjbl_qo, pbadjbl_qo, hneadjbl_qo, cnadjtr_qo, hneadjtr_qo, pbadjtr_qo, cnadjbls_qo,
  pbadjbls_qo, hneadjbls_qo, cnadjtrs_qo, pbadjtrs_qo, hneadjtrs_qo
)

row.names(quantil) <- c(
  "6-CN", "3-PBA", "Creatitine", "4-HNE", "6-CN Baseline", "3-PBA Baseline",
  "4-HNE Baseline", "6-CN Treatment", "3-PBA Treatment", "4-HNE Treatment",
  "6-CN adjusted Baseline", "3-PBA adjusted Baseline", "4-HNE adjusted Baseline",
  "6-CN adjusted Treatment", "3-PBA adjusted Treatment",
  "4-HNE adjusted Treatment", "6-CN log adjusted Baseline centered",
  "3-PBA log adjusted Baseline centered", "4-HNE log adjusted Baseline centered",
  "6-CN log adjusted Treatment centered",
  "3-PBA log adjusted Treatment centered", "4-HNE log adjusted Treatment centered"
)
```

### Quantiles

`r kable(quantil)`

```{r, include=FALSE}
# df_final$phase <- relevel(as.factor(df_final$phase), ref="Morning")

cn_model_days <- lmer(cn_adj_TR_log_S ~ phase + cn_adj_BL_log_S + days + sex
  + phase * days + (1 | record_id), df_final)

tab_model(cn_model_days)

pb_model_days <- lmer(pb_adj_TR_log_S ~ phase + pb_adj_BL_log_S + days + sex
  + phase * days + (1 | record_id), df_final)

tab_model(pb_model_days)

hne_model_days <- lmer(hne_adj_TR_log_S ~ phase + hne_adj_BL_log_S + days + sex
  + phase * days + (1 | record_id), na.omit(df_final))

tab_model(hne_model_days)

hne_pb_model <- lmer(hne_adj_TR_log_S ~ hne_adj_BL_log_S + pb_adj_TR_log_S + days + sex
  + (1 | record_id), na.omit(df_final))

tab_model(hne_pb_model)

hne_cn_model <- lmer(hne_adj_TR_log_S ~ hne_adj_BL_log_S + cn_adj_TR_log_S + days + sex
  + (1 | record_id), na.omit(df_final))

tab_model(hne_cn_model)
```

### 6-CN model
`r tab_model(cn_model_days)`

### 3-PBA model
`r tab_model(pb_model_days)`

### 4-HNE model
`r tab_model(hne_model_days)`

### 4-HNE ~ 3-PBA model
`r tab_model(hne_pb_model)`

### 4-HNE ~ 6-CN model
`r tab_model(hne_cn_model)`

```{r}

df_final_plot <- df_final %>%
  mutate(days = as.factor(days)) %>%
  select(
    days, pb_adj_TR_log_S, pb_adj_BL_log_S, cn_adj_TR_log_S, cn_adj_BL_log_S,
    hne_adj_TR_log_S, hne_adj_BL_log_S, phase
  )
levels(df_final_plot$days) <- c(levels(df_final_plot$days), "3", "6")

df_final_plot_2 <- data.frame(
  as.factor(as.character("3")), NA, NA, NA, NA, NA, NA,
  as.factor(as.character("Evening"))
) %>%
  setNames(names(df_final_plot))

df_final_plot_3 <- data.frame(
  as.factor(as.character("6")), NA, NA, NA, NA, NA, NA,
  as.factor(as.character("Evening"))
) %>%
  setNames(names(df_final_plot))

df_final_plot_4 <- data.frame(
  as.factor(as.character("3")), NA, NA, NA, NA, NA, NA,
  as.factor(as.character("Morning"))
) %>%
  setNames(names(df_final_plot))

df_final_plot_5 <- data.frame(
  as.factor(as.character("6")), NA, NA, NA, NA, NA, NA,
  as.factor(as.character("Morning"))
) %>%
  setNames(names(df_final_plot))

df_final_plot_df <- rbind(
  df_final_plot, df_final_plot_2, df_final_plot_3, df_final_plot_4,
  df_final_plot_5
)

ggplot(df_final_plot_df, aes(x = as.factor(days), y = pb_adj_TR_log_S, fill = phase)) +
  geom_boxplot() +
  geom_jitter() +
  facet_wrap(~phase) +
  xlab("Days after baseline measurement") +
  ylab("3-PBA log adj cent") +
  geom_boxplot(aes(x = as.factor("0=BL"), y = pb_adj_BL_log_S), fill = I("Green")) +
  geom_jitter(aes(x = as.factor("0=BL"), y = pb_adj_BL_log_S)) +
  theme(legend.position = "none")

ggplot(df_final_plot_df, aes(x = as.factor(days), y = cn_adj_TR_log_S, fill = phase)) +
  geom_boxplot() +
  geom_jitter() +
  facet_wrap(~phase) +
  xlab("Days after baseline measurement") +
  ylab("6-CN log adj cent") +
  geom_boxplot(aes(x = as.factor("0=BL"), y = cn_adj_BL_log_S), fill = I("Green")) +
  geom_jitter(aes(x = as.factor("0=BL"), y = cn_adj_BL_log_S)) +
  theme(legend.position = "none")

ggplot(df_final_plot_df, aes(x = as.factor(days), y = hne_adj_TR_log_S, fill = phase)) +
  geom_boxplot() +
  geom_jitter() +
  facet_wrap(~phase) +
  xlab("Days after baseline measurement") +
  ylab("4-HNE log adj cent") +
  geom_boxplot(aes(x = as.factor("0=BL"), y = hne_adj_BL_log_S), fill = I("Green")) +
  geom_jitter(aes(x = as.factor("0=BL"), y = cn_adj_BL_log_S)) +
  theme(legend.position = "none")
```

### 6-CN residual plots

```{r}
plot(cn_model_days, resid(.) ~ fitted(.) | phase)

plot(cn_model_days, resid(.) ~ days)

qqnorm(resid(cn_model_days))
```

### 3-PBA residual plots

```{r }
plot(cn_model_days, resid(.) ~ fitted(.) | phase)

plot(cn_model_days, resid(.) ~ days)

qqnorm(resid(cn_model_days))
```


### 4-HNE residual plots

```{r }
plot(hne_model_days, resid(.) ~ fitted(.) | phase)

plot(hne_model_days, resid(.) ~ days)

qqnorm(resid(hne_model_days))
```


```{r}
ggplot(df_final, aes(y = hne_adj_TR_log_S, x = cn_adj_TR_log_S, color = sex)) +
  geom_point(aes(color = sex)) +
  geom_smooth() +
  theme_classic() +
  ylab("4-HNE log centered adjusted") +
  xlab("6-CN log centered adjusted")

ggplot(df_final, aes(y = hne_adj_TR_log_S, x = pb_adj_TR_log_S, color = sex)) +
  geom_point(aes(color = sex)) +
  geom_smooth() +
  theme_classic() +
  ylab("4-HNE log centered adjusted") +
  xlab("3-PBA log centered adjusted")
```

### Adding the age as a adjustment

```{r}
# df_final$phase <- relevel(as.factor(df_final$phase), ref="Morning")

cn_model_days <- lmer(cn_adj_TR_log_S ~ phase + cn_adj_BL_log_S + days + sex + age
  + phase * days + (1 | record_id), df_final)

tab_model(cn_model_days)

pb_model_days <- lmer(pb_adj_TR_log_S ~ phase + pb_adj_BL_log_S + days + sex + age
  + phase * days + (1 | record_id), df_final)

tab_model(pb_model_days)

hne_model_days <- lmer(hne_adj_TR_log_S ~ phase + hne_adj_BL_log_S + days + sex + age
  + phase * days + (1 | record_id), na.omit(df_final))

tab_model(hne_model_days)

hne_pb_model <- lmer(hne_adj_TR_log_S ~ hne_adj_BL_log_S + pb_adj_TR_log_S + days + sex + age
  + (1 | record_id), na.omit(df_final))

tab_model(hne_pb_model)

hne_cn_model <- lmer(hne_adj_TR_log_S ~ hne_adj_BL_log_S + cn_adj_TR_log_S + days + sex + age
  + (1 | record_id), na.omit(df_final))

tab_model(hne_cn_model)
```

### 6-CN model
`r tab_model(cn_model_days)`

### 3-PBA model
`r tab_model(pb_model_days)`

### 4-HNE model
`r tab_model(hne_model_days)`

### 4-HNE ~ 3-PBA model
`r tab_model(hne_pb_model)`

### 4-HNE ~ 6-CN model
`r tab_model(hne_cn_model)`
